High-resolution ALMA observations of V4046\,Sgr: a circumbinary disc with a thin ring

### 2020-10-15 Draft_V4046Sgr_mnras.tex
Update Draft_V4046Sgr_mnras.tex

% Abstract of the paper
\begin{abstract}
  The nearby V4046\,Sgr spectroscopic binary  hosts a gas-rich disc known for its wide cavity and dusty ring. We present new high resolution ($\sim$20\, mas) ALMA observations of the 1.3\,mm  continuum. The comparison of these  observations, combined with SPHERE-IRDIS polarized  images and a well-sampled spectral energy distribution (SED), against radiative transfer (RT) predictions carried out with the RADMC3D package, allow us to propose a physical model for the source. The new ALMA data reveal a very fine ring at a radius of 13.46$\pm$0.43\,au (Ring13), with a marginally resolved radial width of  4.17$\pm$0.94\,au.  Ring13 is the brightest structure in scattered-light, where it is surrounded by a $\sim$9\,au-wide gap , and it is flanked by a very bright mm outer ring (Ring24) with a sharp inner edge at 24\,au. The steeply decreasing radial slope of Ring24 breaks at $\sim$35\,au into a shallow tail. The RT model requires an inner ring at $\sim$6\,au (Ring6) in small dust grains, hiding under the IRDIS coronagraph, and that surrounds an inner circumbinary disk. Faint mm-continuum coincident with Ring6 is picked up by ALMA, whose morphology  suggests that Ring6  is lopsided or shadowed by the secondary. The previously reported scattered-light shadow of the secondary star is  also reproduced by the RT model.  The surpising  thin Ring13 is nonetheless **10*** times wider than the model scale height, and could thus be long-lived.   The strong near-far disc asymmetry  at 1.65\,$\micron$ points at a very forward-scattering phase function, and requires  grain radii  of no less than  0.4\,$\micron$. 
\end{abstract}

@@

Recent observations of young circumstellar discs have transformed  current knowledge of planet formation, but the focus of resolved imaging with the Atacama Large Millimeter/submillimetre Array (ALMA) or with VLT/SPHERE has  mainly been towards the brighter sources \citep[see][for a review]{Andrews2020arXiv200105007A}. The Disks ARound T Tauri Stars (DARTTS) program was first presented in \citet{Avenhaus_2018} with the aim to image eight stars with the Spectro-Polarimeter High-contrast Exoplanet REsearch (SPHERE). The sample is not biased towards exceptionally bright and large disks, and consists of only solar-mass stars. A second part of the survey increased the number of sources by presenting 21 new images of circumstellar disks \citep{Garufi2020}. The observations revealed diversed structures and morphologies in the scattering surface of these disks. This letter on V4046 Sagittarii (Sgr) is part of our new DARTTS survey with ALMA (DARTTS-A) which will present millimeter observations of nine protoplanetary disks previously imaged in polarized scattered light in DARTTS-S.

@@

For our analysis of the ALMA image we used an image reconstruction strategy, using the non-parametric image synthesis of the \textsc{uvmem} package \citep{2006ApJ...639..951C, 2018A&C....22...16C} it is possible to super-resolve the clean beam, obtaining an effective angular resolution $\sim$3 times finer than  natural weights. The model image is obtained by fitting the data using Chi-squared minimisation, with a measure of regularization when required. Here we adopted a pure $\chi^2$ model image ***sure about this? what is the name of the image you used?****. The resulting \textsc{uvmem} image shown in Fig. \ref{fig:two} exhibits previously unseen substructure of the disc. This image reveals that the disc features two rings of large dust grains with a broad gap between them, i.e. Ring13  at 13\,au and Ring24 starting at 24\,au. The wide and bright  Ring24  reaches its peak intensity at $\sim$30\,au, beyond which it drops steeply before breaking at  $\sim$35\,au into a  shallow tail. While this is the first observation of Ring13,  \citet{Ru_z_Rodr_guez_2019} did anticipate its existence ****HOW? EXPAND ON THIS***

\begin{figure*}
  \includegraphics[width=\textwidth]{hot_two_E.pdf}
  \caption{Comparison of observations and simulated images at 1.65\,$\micron$ and 1.3 mm continuum of the circumbinary disc orbiting V4046\,Sgr. From top to bottom: DPI image and ALMA Band 6 observations; simulated images of the parametric model. Top left panel: SPHERE-IRDIS \textit{H}-band image with a yellow filled circle that illustrate the N\_ALC\_YJH\_S coronagraph (inner working angle $\sim0\farcs12$, or $\sim$8.6\,au at 72.4\,pc). Top right panel: 1.3\,mm continuum \textsc{uvmem} model image. The yellow ellipse shows the size of the natural-weighted beam: $ 0\farcs04 \, \times \, 0\farcs06$. The inset zooms into  the central emission, and the green star marks the star positions. Bottom left panel: 1.65\,$\upmu m$ simulated image of the parametric model. Bottom right panel: 1.3\,mm simulated image of the parametric model. The yellow ellipse shows the size of the  beam: $ 0\farcs04 \, \times \, 0\farcs06$ ***ESTOY CONFUNDIDO POR ESTE BEAM, HAY QUE REPORTAR  1/3 BMAJxBMIN / BPA, INDICANDO 1/3 SEGUIDO POR LOS VALORES EN PESOS NATURALES, Y PLOTEAR EL RESULTADO *****. For all the images in the figure the colour scale is linear.}
  \label{fig:two}
\end{figure*}

Ring13 is surprisingly narrow and seems to be off-centre relative to the GAIA stellar position (at the origin of coordinates in Fig. \ref{fig:two}). We determined the ring's center and orientation by  minimising  the dispersion of the disc radial profiles, from 6\,au to 19\,au, and thus obtained a  PA of 74.6\,$\degr$ ***ADD ERRORES***, an inclination of 33.9\,$\degr$ **ERRORS**** and a centre at $\Delta \alpha = 9\pm0.05$\,mas $\Delta \delta = 0.1\pm0.04$\,mas, relative to the stars.

@@

Repeating the optimization of the disk orientation, but  this time aiming for Ring24  with a radiual domain from 20\,au to 70\,au, we obtained a  PA of 76.8\,$\degr$ ***ERROrSS***, with an inclination of 34.0\,$\degr$ **ERRORS* and a centre at $\Delta \alpha = 19\pm0.04$\,mas $\Delta \delta = 7\pm0.03$\,mas relative to the stars. We see that both Ring13 and Ring24  both share / DO NOT SHARE? the same orientation and center, given the errors, and both are offset from the star. Even though all orientation parameters for Ring13 and Ring24  are consistent within the errors, there  are however some hints for a different orientation, as summarised in Fig. \ref{fig:polarring}, perhaps due to the joint effect of all these small differences. ****CALCULATE THE CHI2 DIFFERNCE BETWEEN THE TWO TRACES? SOME FORM OF STATS WOULD BE GOOD****

@@

\citet{dOrazi} reported a  binary phase in the scattered-light observation around 106\,$\degr$, setting $\phi=$0 from the North and increasing clockwise. Using this measurement, the binary phase was calculated at the time of the ALMA observation at $\phi=$13\,$\degr$. There is no clear signal of any shadow similar to the ones present in the DPI image, this could be explained by innefficient disk cooling compared to the speed of the illumination pattern \citep[][]{Casassus2019MNRAS.486L..58C} ***NO ERA DIFICIL ENCONTRAR ESA REF, PORFA LEER AL MENOS EL ABSTRACT****

@@

The dust begins at 0.3\,au, out of the zone expected to be cleared by dynamical interactions with the central binary \citep{Art_Lu}. For the small dust grains, we propose a three ringed structure. The innermost ring of the model is located from 5 to 8\,au ,this ring is not present in the observations but was added for having a better SED fit. Then follows the two observed rings: the first goes from 12 to 19.5\,au and, with a 1\,au-wide gap between them, the outer ring goes from 20.5\,au and has its maximum density peak at 25\,au. The small-dust density then then fades as the radius increases.

On the other hand, in order to reproduce the two ringed observed morphology of the millimetre continuum emission we required a model that has its large-dust grain population distributed with a wide central cavity (r = 12.3\,au), a narrow 2.6\,au-wide inner ring followed by a gap of 8.2\,au that separates both rings and then an outer ring that goes from 23\,au and has it density peak at 27.3\,au. This last ring has a break between 32 and 34.5\,au and then reaches out to 57\,au.

For the vertical structure, \citet{dOrazi} found flaring angles of $\varphi$ = 6.2$\pm$0.6\,$\degr$ for the inner ring and $\varphi$ = 8.5$\pm$1.0\,$\degr$ for the outer one. For that matter, the model uses two different flaring index $\psi$. The separation between the two values was set at $r$ = 19\,au with values of 0.2 and 0.5 for the inner part and the outer part respectively. Also, the scale height is calculated using $H_o$ = 0.045 and $r_0$ = 19.5\,au.

>>>>

% Abstract of the paper
\begin{abstract}
  The nearby V4046\,Sgr spectroscopic binary hosts a gas-rich disc known for its wide cavity and dusty ring. We present new high resolution ($\sim$20\, mas) ALMA observations of the 1.3\,mm continuum. The comparison of these observations, combined with SPHERE-IRDIS polarized images and a well-sampled spectral energy distribution (SED), against radiative transfer (RT) predictions carried out with the RADMC3D package, allow us to propose a physical model for the source. The new ALMA data reveal a very fine ring at a radius of 13.46$\pm$0.43\,au (Ring13), with a marginally resolved radial width of 4.17$\pm$0.94\,au. Ring13 is the brightest structure in scattered-light, where it is surrounded by a $\sim$9\,au-wide gap , and it is flanked by a very bright mm outer ring (Ring24) with a sharp inner edge at 24\,au. The steeply decreasing radial slope of Ring24 breaks at $\sim$35\,au into a shallow tail. The RT model requires an inner ring at $\sim$6\,au (Ring6) in small dust grains, hiding under the IRDIS coronagraph, and that surrounds an inner circumbinary disk. Faint mm-continuum coincident with Ring6 is picked up by ALMA, whose morphology suggests that Ring6 is lopsided or shadowed by the secondary. The previously reported scattered-light shadow of the secondary star is also reproduced by the RT model. The surpising thin Ring13 is nonetheless **15** times wider than the model scale height, and could thus be long-lived. The strong near-far disc asymmetry at 1.65\,$\micron$ points at a very forward-scattering phase function, and requires grain radii of no less than 0.4\,$\micron$. 
\end{abstract}

@@

Recent observations of young circumstellar discs have transformed current knowledge of planet formation, but the focus of resolved imaging with the Atacama Large Millimeter/submillimetre Array (ALMA) or with VLT/SPHERE has mainly been towards the brighter sources \citep{Andrews2020arXiv200105007A}. The Disks ARound T Tauri Stars (DARTTS) program was first presented in \citet{Avenhaus_2018} with the aim to image eight stars with the Spectro-Polarimeter High-contrast Exoplanet REsearch (SPHERE). The sample is not biased towards exceptionally bright and large disks, and consists of only solar-mass stars. A second part of the survey increased the number of sources by presenting 21 new images of circumstellar disks \citep{Garufi2020}. The observations revealed diversed structures and morphologies in the scattering surface of these disks. This letter on V4046 Sagittarii (Sgr) is part of our new DARTTS survey with ALMA (DARTTS-A) which will present millimeter observations of nine protoplanetary disks previously imaged in polarized scattered light in DARTTS-S.

@@

For our analysis of the ALMA image we used an image reconstruction strategy, using the non-parametric image synthesis of the \textsc{uvmem} package \citep{2006ApJ...639..951C, 2018A&C....22...16C} it is possible to super-resolve the clean beam, obtaining an effective angular resolution $\sim$3 times finer than natural weights. The model image is obtained by fitting the data using Chi-squared minimisation, with a measure of regularization when required. Here we adopted a pure $\chi^2$ model image. The resulting \textsc{uvmem} image shown in Fig. \ref{fig:two} exhibits previously unseen substructure of the disc. This image reveals that the disc features two rings of large dust grains with a broad gap between them, i.e. Ring13 at 13\,au and Ring24 starting at 24\,au. The wide and bright Ring24 reaches its peak intensity at $\sim$30\,au, beyond which it drops steeply before breaking at $\sim$35\,au into a shallow tail. While this is the first observation of Ring13, \citet{Ru_z_Rodr_guez_2019} did anticipate its existence as their ALMA continuum image showed a distinct excess between
$\sim$10 and 17\,au.

\begin{figure*}
  \includegraphics[width=\textwidth]{hot_two_E.pdf}
  \caption{Comparison of observations and simulated images at 1.65\,$\micron$ and 1.3 mm continuum of the circumbinary disc orbiting V4046\,Sgr. From top to bottom: DPI image and ALMA Band 6 observations; simulated images of the parametric model. Top left panel: SPHERE-IRDIS \textit{H}-band image with a yellow filled circle that illustrate the N\_ALC\_YJH\_S coronagraph (inner working angle $\sim0\farcs12$, or $\sim$8.6\,au at 72.4\,pc). Top right panel: 1.3\,mm continuum \textsc{uvmem} model image. The yellow ellipse shows the size of the natural-weighted beam: $ 0\farcs04 \, \times \, 0\farcs06$. The inset zooms into the central emission, and the green star marks the star positions. Bottom left panel: 1.65\,$\upmu m$ simulated image of the parametric model. Bottom right panel: 1.3\,mm simulated image of the parametric model. The yellow ellipse shows the size of the beam: $0\farcs04 \, \times \, 0\farcs06$ ***ESTOY CONFUNDIDO POR ESTE BEAM, HAY QUE REPORTAR 1/3 BMAJxBMIN / BPA, INDICANDO 1/3 SEGUIDO POR LOS VALORES EN PESOS NATURALES, Y PLOTEAR EL RESULTADO *****. For all the images in the figure the colour scale is linear.}
  \label{fig:two}
\end{figure*}

Ring13 is surprisingly narrow and seems to be off-centre relative to the GAIA stellar position (at the origin of coordinates in Fig. \ref{fig:two}). We determined the ring's centre and orientation by minimising the dispersion of the disc radial profiles, from 6\,au to 19\,au, and thus obtained a PA of 74.64$\pm$0.08\,$\degr$, an inclination of 33.90$\pm$0.04\,$\degr$ and a centre at $\Delta \alpha = 9\pm0.05$\,mas $\Delta \delta = 0.1\pm0.04$\,mas, relative to the stars.

@@

Repeating the optimization of the disk orientation, but this time aiming for Ring24 with a radial domain from 20\,au to 70\,au, we obtained a PA of 76.78$\pm$0.03\,$\degr$, with an inclination of 34.01$\pm$0.01\,$\degr$ and a centre at $\Delta \alpha = 19\pm0.04$\,mas $\Delta \delta = 7\pm0.03$\,mas relative to the stars. We see that both Ring13 and Ring24 both share / DO NOT SHARE? the same orientation and centre, given the errors, and both are offset from the star. Even though all orientation parameters for Ring13 and Ring24 are consistent within the errors, there are however some hints for a different orientation, as summarised in Fig. \ref{fig:polarring}, perhaps due to the joint effect of all these small differences. ****CALCULATE THE CHI2 DIFFERENCE BETWEEN THE TWO TRACES? SOME FORM OF STATS WOULD BE GOOD****

@@

\citet{dOrazi} reported a binary phase in the scattered-light observation around 106\,$\degr$, setting $\phi=$0 from the North and increasing clockwise. Using this measurement, the binary phase was calculated at the time of the ALMA observation at $\phi=$13\,$\degr$. There is no clear signal of any shadow similar to the ones present in the DPI image, this could be explained by inefficient disk cooling compared to the speed of the illumination pattern \citep{Casassus2019MNRAS.486L..58C}.

@@

The dust begins at 0.3\,au, out of the zone expected to be cleared by dynamical interactions with the central binary \citep{Art_Lu}. For the small dust grains, we propose a three ringed structure. The innermost ring of the model is located from 5 to 8\,au (Ring6),this ring is not present in the observations but was added for having a better SED fit. Then follows the two observed rings: the first goes from 12 to 19.5\,au (Ring13) and, with a 1\,au-wide gap between them, the outer ring (Ring24) goes from 20.5\,au and has its maximum density peak at 25\,au. The small-dust density then then fades as the radius increases.

On the other hand, in order to reproduce the two ringed observed morphology of the millimetre continuum emission we required a model that has its large-dust grain population distributed with a wide central cavity (r = 12.3\,au), a narrow 2.6\,au-wide inner ring (Ring13) followed by a gap of 8.2\,au that separates both rings and then an outer ring (Ring24) that goes from 23\,au and has it density peak at 27.3\,au. This last ring has a break between 32 and 34.5\,au and then reaches out to 57\,au.

For the vertical structure, \citet{dOrazi} found flaring angles of $\varphi$ = 6.2$\pm$0.6\,$\degr$ for the inner ring and $\varphi$ = 8.5$\pm$1.0\,$\degr$ for the outer one. For that matter, the model uses two different flaring index $\psi$. The separation between the two values was set at $r$ = 19\,au with values of 0.2 for the Ring6 and Ring13 and 0.5 for Ring24. Also, the scale height is calculated using $H_o$ = 0.045 and $r_0$ = 19.5\,au.
